---
title: "rWhatsApp"
format:
  html:
    embed-resources: true
  gfm: default
---

# rwhatsapp: Import and Handling for WhatsApp Chat Logs


## Description
`rwhatsapp` is a straightforward, robust R package for importing and parsing exported WhatsApp chat logs (Android/iOS) into tidy tabular data. It ingests `.txt` and `.zip` history files across locales/devices and exposes message text, timestamps, authors, and emojis for downstream analysis.

## Use Cases
- Import exported WhatsApp chats to study interpersonal communication
- Describe conversations (e.g., messages per day; messages per author) to study activity and participation.
- Explore emoji usage patterns and look up emoji names to understand affective dynamics.
- Conduct token-based analyses (e.g., most-used words, tf–idf per author) with standard text-mining workflows.
- Teaching: demonstrate tidy text pipelines on personal communication data.

## Input Data
- **Accepted inputs:** WhatsApp chat exports in plain text (`.txt`) or compressed archives (`.zip`) created via the app’s export function.
- **Locales/devices:** Designed to handle multiple locales and both Android/iOS.
- **Example files that come with the package:** sample.txt[`system.file("extdata", "sample.txt", package = "rwhatsapp")`]

### Typical Structure
Lines with timestamps, author handles, message text; media as placeholders. If you extract it and open the .txt file it should look like this:

[06/03/25, 2:27:42] Markus: I just need a quick coffee, and then I'll be back to normal. Hahaha

[06/03/25, 2:27:47] Angela: My day has been totally random, but I'll be right back as well!

## Output Data
`rwa_read()` returns a tibble:
- `time`, `author`, `text`, `source`, `emoji`, `emoji_name`
```{r}
library("rwhatsapp")
history <- system.file("extdata", "sample.txt", package = "rwhatsapp")
chat <- rwa_read(history)
head(chat)
```

## Hardware Requirements
- **R:** ≥ 3.5.0
- **Resources:** Small-to-moderate for typical chats.

## Environment Setup
```{r, eval=FALSE}
install.packages("rwhatsapp")
```
OR
```{r, eval=FALSE}
remotes::install_github("JBGruber/rwhatsapp")
```

### Loading Dependencies:
RWhatsApp needs a number of other packages to be installed to ensure functionality. Always make sure to load these common packages into your environment when using RWhatsapp:
```{r}
library("rwhatsapp")
library("dplyr")
library("ggplot2")
library("lubridate")
library("tidyr")
library("tidytext")
library("stopwords")
```

## How to Use

In the following example we will show you how to use the main features of the package. You may use the example data or are welcome to load your own chatlogs into the environment.

### Load sample data and read into R
```{r}
history <- system.file("extdata", "sample.txt", package = "rwhatsapp")
chat <- rwa_read(history)
chat
```

### Loading own data
```{r, eval=FALSE}
chat <- rwa_read("/path/to/your/WhatsAppChat.txt")
```

### Remove system messages
In this step we are removing any messages, that are not sent from one of the two human authors in the chat:
```{r}
chat <- subset(chat, !is.na(author))
```

### Messages per day
We are now able to extract i.E. how many messages were sent per day or on a specific day by both authors:
```{r}
theme_set(theme_minimal())

chat |>
  mutate(day = date(time)) |>
  count(day) |>
  ggplot(aes(day, n)) +
  geom_col() +
  labs(
    title = "Messages per Day",
    x = "Date",
    y = "Number of Messages"
  ) +
  scale_x_date(date_labels = "%d.%m.%Y")
```

### Messages per author
If we want to split the message count, we can use the following feature to check how many messages were sent by each author:
```{r}
chat |>
  count(author) |>
  ggplot(aes(x = reorder(author, n), y = n)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Messages per Author",
    x = NULL,
    y = "Number of Messages"
  ) +
  theme_minimal()
```

### Emoji exploration
Sometimes it is of interest, which and how often emojis are used in a given chat. By using the following code chunk we can examine this further:
```{r}
chat |>
  unnest(emoji) |>
  count(author, emoji, sort = TRUE) |>
  rename(
    Author = author,
    Emoji = emoji,
    Frequency = n
  )
```

### Token-based exploration
We may also try to see how often a given author used a specific word/wording. By using the following code chunk we can examine this further:
```{r}
chat |>
  tidytext::unnest_tokens(input = text, output = word) |>
  count(author, word, sort = TRUE) |>
  rename(
    Author = author,
    Word = word,
    Frequency = n
  ) |>
  select(Author, Word, Frequency)
```

### Deep Dive
After you are familiar with the package and its basic functionalities you are able to combine the diverse feature together to undergo deeper analysis. The following example will not work well with the example data, but works great if you have loaded your own WhatsApp-Data. If you are unable to do so, we have provided a larger example dataset for this task.

Another way to determine favourite words is to calculate the term frequency–inverse document frequency (tf–idf). Basically, what the measure does, in this case, is to find words that are common within the messages of one author but uncommon in the rest of the messages.
```{r}
# First we define stopwords to be removed from the analysis. You may need to add further words to this list while investigating the dataset.

to_remove <- c(stopwords(language = "en"),
               "example")

# Then we load our own dataframe (or leave it as is to use our larger example dataset)
chat <- rwa_read("ExampleChat.txt")
chat <- subset(chat, !is.na(author))

chat %>%
  unnest_tokens(input = text,
                output = word) %>%
  select(word, author) %>%
  filter(!word %in% to_remove) %>%
  mutate(word = gsub(".com", "", word)) %>%
  mutate(word = gsub("^gag", "9gag", word)) %>%
  count(author, word, sort = TRUE) %>%
  bind_tf_idf(term = word, document = author, n = n) %>%
  filter(n > 10) %>%
  group_by(author) %>%
  top_n(n = 6, tf_idf) %>%
  ggplot(aes(x = reorder_within(word, n, author), y = n, fill = author)) +
  geom_col(show.legend = FALSE) +
  ylab("") +
  xlab("") +
  coord_flip() +
  facet_wrap(~author, ncol = 2, scales = "free_y") +
  scale_x_reordered() +
  ggtitle("Important words using tf–idf by author")

```
For further deep dives and more in depth tutorials visit [the official documentation page](https://cran.r-project.org/web/packages/rwhatsapp/vignettes/Text_Analysis_using_WhatsApp_data.html).

## Technical Details
See the [CRAN REPOSITORY](https://cran.r-project.org/package=rwhatsapp) for technical information.

## References
- Gruber, J. (2022). *rwhatsapp: Import and handling for “WhatsApp” chat logs* (Version 0.2.4). CRAN. [https://doi.org/10.32614/CRAN.package.rwhatsapp](https://doi.org/10.32614/CRAN.package.rwhatsapp)

## Disclaimer
This method is intended for academic research; ensure legality and ethics when processing chat data.

## Contact Details
- Issues: <https://github.com/JBGruber/rwhatsapp/issues>
- Contact: Johannes Gruber <JohannesB.Gruber@gmail.com>
